Objective-first optimization for nanophotonics

### 2012-05-31 presentations/meta_book/ob1.tex
New pictures added.

\subsection{Typical design formulation}
The typical, and most straightforward formulation
    of the design problem is
\BA \minimize{x,p} f(x) \\
    \subto A(p)x - b(p) = 0, \label{eq:typ:eqcon} \EA
    which states that we would like to vary $x$ and $p$ simultaneously
    in order to decrease $f(x)$
    while always satisfying physics (the electromagnetic wave equation).

Such a formulation, if solved using a steepest-descent method,
    is the well-known adjoint optimization method. % Ref.

\subsection{Objective-first design formulation}
In contrast, the objective-first formulation is 
\BA \minimize{x,p} \| A(p) x - b(p) \|^2 \\
    \subto f(x) = f_\text{ideal}, \EA
    where the roles of the wave equation and the design objective are switched.

This is a bi-convex problem, which we solve using an alternating directions method.

\subsection{Field sub-problem}
\BA \minimize{x} \| A(p) x - b(p) \|^2 \\
    \subto f(x) = f_\text{ideal} \EA

\subsection{Structure sub-problem}
\BA \minimize{p} \| B(x) p - d(x) \|^2 \\
    \subto 0 \le p \le 1 \EA

>>>>

\subsection{Typical design formulation}
The typical, and most straightforward formulation
    of the design problem is
\BA \minimize{x,p} f(x) \\
    \subto A(p)x - b(p) = 0, \label{eq:typ:eqcon} \EA
    which states that we would like to vary $x$ and $p$ simultaneously
    in order to decrease $f(x)$
    while always satisfying physics (the electromagnetic wave equation).
Such a formulation, if solved using a steepest-descent method,
    is the well-known adjoint optimization method. % Ref.

\subsection{Objective-first design formulation}
In contrast, the objective-first formulation switches
    the roles of the wave equation and the design objective
    with one another,
\BA \minimize{x,p} \| A(p) x - b(p) \|^2 \label{eq:ob1:1} \\
    \subto f(x) = f_\text{ideal}. \label{eq:ob1:2} \EA

This first means that, as seen from \ER{ob1:1},
    we allow for non-zero residual in the electromagnetic wave equation.
This literally means that we allow for \emph{non-physical} $x$ and $p$,
    since $A(p) x - b(p) \ne 0$ is now allowed.
For this reason, we denote $A(p) x - b(p)$ the \emph{physics residual}.

Secondly, we see from \ER{ob1:2} that we always force the device
    to exhibit ideal performance,
    even if doing so means breaking the laws of physics.
As such, our strategy will be to vary $x$ and $p$
    in order to decrease the physics residual \ER{ob1:1} to zero,
    while always maintaining ideal performance.

Our initial motivation for doing this 
    was that it allowed us to solve for $x$ and $p$ separately,
    as will be outlined below,
    and that always forcing ideal performance
    might provide a mechanism to ``override'' local optima 
    in the optimization process.

To this end we have found that such a strategy
    actually allows us to design very unintuitive devices
    which exhibit very good performance,
    even when starting from completely non-functional initial guesses.
Furthermore, we have found this to be true
    even true when the physics residual is never brought to exactly zero.

From a numerical standpoint, 
    although the objective-first formulation is still non-convex 
    in its original form,
    the bi-linearity of the physics residual term allows us to 
    naturally break the original problem into two sub-problems
    which we outline below.

Lastly, we add an additional constraint to the original formulation,
    which is to set hard-limits on the allowable values of $p$,
    namely $p_0 \le p \le p_1$.
This is actually a relaxation of the ideal constraint,
    which would be to allow $p$ to only have discrete values,
    $p \in p_0, p_1$,
    but such a constraint would be essentially force us to only
    be able to perform brute force trial-and-error.

Our objective-first formulation is thus,
\BA \minimize{x,p} \| A(p) x - b(p) \|^2 \notag \\
    \subto f(x) = f_\text{ideal} \label{eq:ob1} \\
        & p_0 \le p \le p_1. \notag \EA

\subsection{Field sub-problem}
Since the objective-first problem in its original form is still non-convex,
    we break it down into two convex sub-problems.
The first of these is the field sub-problem, 
    which simply involves fixing $p$ and independently optimizing $x$,
\BA \minimize{x} \| A(p) x - b(p) \|^2 \label{eq:Fsub} \\
    \subto f(x) = f_\text{ideal}. \notag \EA
This problem is convex, and actually quadratic,
    which means that it can even be solved in the same way 
    as a simple least-squares problem.

\subsection{Structure sub-problem}
The second sub-problem is formulated by fixing $x$ and
    independently optimizing $p$.
At the same time, we use the bi-linearity property
    of the physics residual from \ER{bilinear}
    to rewrite the problem in a way that makes
    its convexity explicit,
\BA \minimize{p} \| B(x) p - d(x) \|^2 \label{eq:Ssub} \\
    \subto p_0 \le p \le p_1. \notag \EA

The structure sub-problem is also convex, but not quadratic.
However, use of the CVX package still allows us to obtain the result
    quickly and reliably.

\subsection{Alternating directions algorithm}
We use a simple alternating directions scheme 
    to peice together \ER{Fsub} and \ER{Ssub},
    which is to say that we simply
    alternately solve each and continue until we reach some stopping point,
    normally measured by how much the physics residual has decreased.

The advantage of such the alternating directions method
    is that the physics residual is guaranteed to
    monotonically decrease with every iteration,
    which is useful in that no safeguards
    are needed to guard against ``rogue'' steps
    in the optimization procedure.
Note that this robustness stems from the fact that,
    among other things,
    each sub-problem does not rely on previous values of 
    the variable which is being optimized,
    but only on the variable which is held constant.

The disadvantage of such a simple scheme is that 
    the convergence is quite slow,
    although we have found it to be sufficient in our cases.
See ref % cite!
    for related methods that exhibit far better convergence.

### 2012-05-31 presentations/meta_book/wave_eq.tex
New pictures added.

We start by understanding the underlying physical equation.
We extend it beyond its usual use in simulation, 
    and use it for design.

@@

\subsection{Bi-linearity of the wave equation}
Although additional mathematical machinery must still be added
    in order to get a useful design tool,
    we have really shown so far is that the wave equation is 
    separately linear or \emph{bi-linear} in $x$ and $p$.
Namely that,
\BE A(p)x-b(p) = B(x)p - d(x). \EE
In other words, fixing $p$ makes solving the wave equation for $x$
    a linear problem, and vice versa.
Note that the joint problem,
    where both $x$ and $p$ are allowed to vary,
    is not linear.

>>>>

@@

\subsection{Bi-linearity of the wave equation}
Although additional mathematical machinery must still be added
    in order to get a useful design tool,
    we have really shown so far is that the wave equation is 
    separately linear or \emph{bi-linear} in $x$ and $p$.
Namely that,
\BE A(p)x-b(p) = B(x)p - d(x). \label{eq:bilinear} \EE
In other words, fixing $p$ makes solving the wave equation for $x$
    a linear problem, and vice versa.
Note that the joint problem,
    where both $x$ and $p$ are allowed to vary,
    is not linear.
